---
title: "Week 9 - Multiple Linear Regression"
output: html_document
---

### Load Packages:

```{r}
library(tidyverse)
library(openintro)
library(GGally)
```

### The Data

The data were gathered from end of semester student evaluations for a large sample of professors from the University of Texas at Austin. In addition, six students rated the professors’ physical appearance. The result is a data frame where each row contains a different course and columns represent variables about the courses and professors. It’s called evals.

```{r}
view(evals)
```


```{r}
glimpse(evals)
```

### Exploring the data:

#### Exercise 1:

Is this an observational study or an experiment? The original research question posed in the paper is whether beauty leads directly to the differences in course evaluations. Given the study design, is it possible to answer this question as it is phrased? If not, rephrase the question.

Answer:Yes this is an observational study since there is no experiment designed to collect data. And yes, it is possible to answerthe question that there may be correlation between the two variables and not the causation.

#### Exercise 2:

Describe the distribution of score. Is the distribution skewed? What does that tell you about how students rate courses? Is this what you expected to see? Why, or why not?

Answer:The distribution is left skewed, which means that the mean is less than the median.There are more positive than negative evaluation for the teachers.Yes this is what I would expect, because there are mostly fewer teachers whose rating goes negative.

```{r}
hist(evals$score)
```

#### Exercise 3:

Excluding score, select two other variables and describe their relationship with each other using an appropriate visualization.

Answer:The data points are scattered all over, not a relationship between the teacher’s age and beauty score.

```{r}
#boxplot(evals$bty_avg ~ evals$age)

#fit simple linear regression model
model <- lm(evals$age ~ evals$bty_avg, data = evals)

#create scatterplot of data
plot(evals$bty_avg, evals$age)

#add fitted regression line
abline(model)
```

### Simple linear regression:
Answer:The fundamental phenomenon suggested by the study is that better looking teachers are evaluated more favorably. Let’s create a scatterplot to see if this appears to be the case:

```{r}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_point()
```

Before we draw conclusions about the trend, compare the number of observations in the data frame with the approximate number of points on the scatterplot. Is anything awry?
There seem to be more observations than the approximate number of points on the scatterplot.

```{r}
nrow(evals)
```

###Exercise 4: 

Replot the scatterplot, but this time use geom_jitter as your layer. What was misleading about the initial scatterplot?

Answer: It is not showing the relationship between beauty average and score for the teacher because of overlapping scores
```{r}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter()
```

###Exercise 5: 

Let’s see if the apparent trend in the plot is something more than natural variation. Fit a linear model called m_bty to predict average professor score by average beauty rating and add the line to your plot using abline(m_bty). Write out the equation for the linear model and interpret the slope. Is average beauty score a statistically significant predictor? Does it appear to be a practically significant predictor?

```{r}
m_bty <- lm(evals$score~evals$bty_avg)
plot(jitter(evals$score) ~ jitter(evals$bty_avg))
abline(m_bty)
```
The blue line is the model. The shaded gray area around the line tells you about the variability you might expect in your predictions. To turn that off, use se = FALSE.


```{r}
summary(m_bty)
```
Answer:Answer: Investigating the relationships between score and the bty_avg variable:

              y^=3.88034+0.06664∗btyavg


This appears to be a practically significant predictor; as many points are nearer (above or below) to the line.


###Exercise 6: 

Use residual plots to evaluate whether the conditions of least squares regression are reasonable. Provide plots and comments for each one (see the Simple Regression Lab for a reminder of how to make these).

### adds a horizontal dashed line at y = 0
```{r}
plot(m_bty$residuals ~ evals$bty_avg)
abline(h = 0, lty = 3) 
```
```{r}
hist(m_bty$residuals)
```

### adds diagonal line to the normal prob plot

```{r}
qqnorm(m_bty$residuals)
qqline(m_bty$residuals)
```

Answer: According to the histogram plot the residuals appear to be skewed right. But we can say that the linear model for relationship between variable score and bty_avg is somewhat reliable.

### Multiple linear regression:

The data set contains several variables on the beauty score of the professor: individual ratings from each of the six students who were asked to score the physical appearance of the professors and the average of these six scores. Let’s take a look at the relationship between one of these scores and the average beauty score.

```{r}
plot(evals$bty_avg ~ evals$bty_f1lower)
```
As expected, the relationship is quite strong—after all, the average score is calculated using the individual scores. You can actually look at the relationships between all beauty variables (columns 13 through 19) using the following command:

```{r}
cor(evals$bty_avg, evals$bty_f1lower)
```
Another way of calculation
```{r}
evals %>% 
  summarise(cor(bty_avg, bty_f1lower))
```
```{r}
plot(evals[,13:19])
```

```{r}
evals %>%
  select(contains("bty")) %>%
  ggpairs()
```
These variables are collinear (correlated), and adding more than one of these variables to the model would not add much value to the model. In this application and with these highly-correlated predictors, it is reasonable to use the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score after you’ve accounted for the professor’s gender, you can add the gender term into the model.

```{r}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

###Exercise 7: 

P-values and parameter estimates should only be trusted if the conditions for the regression are reasonable. Verify that the conditions for this model are reasonable using diagnostic plots.

### adds a horizontal dashed line at y = 0
```{r}
plot(m_bty_gen$residuals ~ evals$bty_avg + evals$gender)
abline(h = 0, lty = 3) 
```

```{r}
hist(m_bty_gen$residuals)
```
### adds diagonal line to the normal prob plot

Answer: Using diagnostic plots, we can conclude that the conditions for this model are reasonable. The variables are linearly related to the outcome.
 
```{r}
qqnorm(m_bty_gen$residuals)
qqline(m_bty_gen$residuals) 
```

###Exercise 8: 

Is bty_avg still a significant predictor of score? Has the addition of gender to the model changed the parameter estimate for bty_avg?

Answer: Yes, bty_avg still a significant predictor of score; but not by much.


Note that the estimate for gender is now called gendermale. You’ll see this name change whenever you introduce a categorical variable. The reason is that R recodes gender from having the values of male and female to being an indicator variable called gendermale that takes a value of 0 for female professors and a value of 1 for male professors. (Such variables are often referred to as “dummy” variables.)

As a result, for female professors, the parameter estimate is multiplied by zero, leaving the intercept and slope form familiar from simple regression.

scoreˆ=β^0+β^1×bty_avg+β^2×(0)

      =β^0+β^1×bty_avg
      
We can plot this line and the line corresponding to males with the following custom function.

```{r}
# multiLines(m_bty_gen)
```

###Exercise 9: 

What is the equation of the line corresponding to males? (Hint: For males, the parameter estimate is multiplied by 1.) For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

Answer: the equation of the line corresponding to males is:

y^=3.74734+0.17239∗gendermale
.

For two professors who received the same beauty rating, the MALE gender tends to have the higher course evaluation score.


The decision to call the indicator variable gendermale instead ofgenderfemale has no deeper meaning. R simply codes the category that comes first alphabetically as a 0. (You can change the reference level of a categorical variable, which is the level that is coded as a 0, using therelevel function. Use ?relevel to learn more.)

###Exercise 10: 

Create a new model called m_bty_rank with gender removed and rank added in. How does R appear to handle categorical variables that have more than two levels? Note that the rank variable has three levels: teaching, tenure track, tenured.

```{r}
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)
```
R has created dummy variables for categorical variables with more than two levels. In this case - ranktenure track, and ranktenured.


The interpretation of the coefficients in multiple regression is slightly different from that of simple regression. The estimate for bty_avg reflects how much higher a group of professors is expected to score if they have a beauty rating that is one point higher while holding all other variables constant. In this case, that translates into considering only professors of the same rank with bty_avg scores that are one point apart.

### The search for the best model:

We will start with a full model that predicts professor score based on rank, ethnicity, gender, language of the university where they got their degree, age, proportion of students that filled out evaluations, class size, course level, number of professors, number of credits, average beauty rating, outfit, and picture color.

###Exercise 11: 

Which variable would you expect to have the highest p-value in this model? Why? Hint: Think about which variable would you expect to not have any association with the professor score.

Answer: The vairable cls_profs, number of professors teaching sections in course in sample; has no association and thus expect it to have highest p-value.

Let’s run the model…

```{r}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```
###Exercise 12: 

Check your suspicions from the previous exercise. Include the model output in your response..

Answer: We can see from above that the variable cls_profs thus have highest estiamte.

###Exercise 13: 

Interpret the coefficient associated with the ethnicity variable.

Answer: Non-minority professors will see 0.12 points higher than minority professors.

###Exercise 14:

Drop the variable with the highest p-value and re-fit the model. Did the coefficients and significance of the other explanatory variables change? (One of the things that makes multiple regression interesting is that coefficient estimates depend on the other variables that are included in the model.) If not, what does this say about whether or not the dropped variable was collinear with the other explanatory variables?

```{r}
m_full.new <- lm(score ~ ethnicity + gender + language + age 
                 + cls_perc_eval + cls_credits + bty_avg 
                 + pic_color, data = evals)

summary(m_full.new)
```

###Exercise 15:

Using backward-selection and p-value as the selection criterion, determine the best model. You do not need to show all steps in your answer, just the output for the final model. Also, write out the linear model for predicting score based on the final model you settle on

```{r}
m_full.final <- lm(score ~  ethnicity + gender + language + age 
                   + cls_perc_eval + cls_credits + bty_avg 
                   + pic_color, data = evals)

summary(m_full.final)
```

###Exercise 16:

Verify that the conditions for this model are reasonable using diagnostic plots

```{r}
plot(m_full.final$residuals ~ evals$bty_avg)
abline(h = 0, lty = 3)
```

```{r}
hist(m_full.final$residuals)
```

```{r}
qqnorm(m_full.final$residuals)
qqline(m_full.final$residuals)
```

Answer: Residual are nearly normal; The hist plot seems normal although leftward skewed.

###Exercise 17:

The original paper describes how these data were gathered by taking a sample of professors from the University of Texas at Austin and including all courses that they have taught. Considering that each row represents a course, could this new information have an impact on any of the conditions of linear regression?

Answer: Yes, it does have an impact on any of the conditions of linear regression since independence will be violated.


###Exercise 18:

Based on your final model, describe the characteristics of a professor and course at University of Texas at Austin that would be associated with a high evaluation score.

Answer: Based on the formulae on #15; we can say a Non-Minority Male,
who is Non-English Speaking, has one credit per class, and has higher percent of students evaluating. 

###Exercise 19:

Would you be comfortable generalizing your conclusions to apply to professors generally (at any university)? Why or why not?

Answer: No, we would want to run several more studies. The given data set would not be a good representation of any other universities student profiles.