---
title: 'tidyverse: Worked Example'
author: "Mubashira Qari"
date: "April 22, 2022"
output: html_document
---

## Readr -- Reading Files

### The first package of TidyVerse to read the file is "Readr" It is a fast way to read rectangular data from delimited files, such as comma-separated values (CSV) and tab-separated values (TSV).

###The easiest way to get readr is to install the whole tidyverse:

readr supports the following file formats with these read_*() functions:

read_csv(): comma-separated values (CSV) files
read_tsv(): tab-separated values (TSV) files
read_delim(): delimited files (CSV and TSV are important special cases)
read_fwf(): fixed-width files
read_table(): whitespace-separated files
read_log(): web log files

```{r}
library(tidyverse)
library(readr)
library(dplyr)
```

### Fetching the files from relational dataset on GitHub
```{r}
college_major <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/majors-list.csv", na="")

view(college_major)
```

### Deplyr -- Rename Columns

For rename(): <tidy-select> Use new_name = old_name to rename selected variables.

```{r}
college_major <- as_tibble(college_major) # so it prints a little nicer
```


```{r}

college_major <- rename(college_major, Major_Id = FOD1P)

head(college_major)
```



```{r}
all_ages <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/all-ages.csv", na="")

view(all_ages)
```
### To generate query

```{r}
summary <- all_ages %>% 
  group_by(Major_category) %>% 
  summarise(Employed = mean(Employed, na.rm = TRUE)) %>% 
  arrange(desc(Employed))

summary
```


```{r}

```


```{r}
grad_students <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/grad-students.csv", na="")

view(grad_students)
```


```{r}
women_stem <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/grad-students.csv", na="")

view(women_stem)
```

```{r}
recent_grads <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/recent-grads.csv", na="")

view(recent_grads)
```

###

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

