---
title: 'TidyVerse: Worked Example'
author: "Mubashira Qari"
date: "April 22, 2022"
output: html_document
---

### Objective: Create an Example:  
I am using couple of TidyVerse packages here, and college 'major_list' dataset from fivethirtyeight.com , and creating a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with my selected dataset.


### Readr -- Reading Files

The first package of TidyVerse to read the file is "Readr" It is a fast way to read rectangular data from delimited files, such as comma-separated values (CSV) and tab-separated values (TSV).

### The easiest way to get readr is to install the whole tidyverse:

readr supports the following file formats with these read_*() functions:

read_csv(): comma-separated values (CSV) files
read_tsv(): tab-separated values (TSV) files
read_delim(): delimited files (CSV and TSV are important special cases)
read_fwf(): fixed-width files
read_table(): whitespace-separated files
read_log(): web log files

```{r}
library(tidyverse)
library(readr)
library(dplyr)
```

### Fetching the majors-list dataset from the relational dataset on GitHub

```{r}
college_major <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/majors-list.csv", na="")

head(college_major)
```

### Using package 'Deplyr' -- to Rename Columns

For rename(): <tidy-select> Use new_name = old_name to rename selected variables.

```{r}
college_major <- as_tibble(college_major) # so it prints a little nicer
```


```{r}

college_major <- rename(college_major, Major_Id = FOD1P)

head(college_major)
```

### Reading 'all-ages' dataset

```{r}
all_ages <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/all-ages.csv", na="")

head(all_ages)
```
### This 'pipe operator' replicate SQL query 

```{r}
summary <- all_ages %>% 
  group_by(Major_category) %>% 
  summarise(Employed = mean(Employed, na.rm = TRUE)) %>% 
  arrange(desc(Employed))

summary
```

### Using the 'ggplot' for visualizing which category has most most employed

```{r}
ggplot(summary, aes(x = Employed, y = Major_category, fill = Major_category)) +
  geom_col(position = "dodge")
```

### Reading the 'grad-students' dataset

```{r}
grad_students <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/grad-students.csv", na="")

head(grad_students)
```

### Mutating joins:

Mutating joins allow you to combine variables from multiple tables.

```{r}
merge_df <- all_ages %>% left_join(grad_students, by = "Major_code")

merge_df
```

### Remove multiple Column in R using dplyr (by name and index) -- TidyVerse

```{r}

merge_df <- select(merge_df, -c(Major_category.y, Major.y))

head(merge_df)
```

### Change Name of Columns in R with dplyr rename()

```{r}
merge_df <- merge_df %>%
  rename(Major_category = Major_category.x,
         Major = Major.x
         )

head(merge_df)
```

### Reading the 'grad-students' dataset

```{r}
women_stem <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/grad-students.csv", na="")

head(women_stem)
```

### Reading the 'recent-grads' dataset

```{r}
recent_grads <- read_csv("https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/tidyverse_assignment/recent-grads.csv", na="")

head(recent_grads)
```

### Merging merge_df and recent_grads

```{r}
merge_df <- merge_df %>% left_join(recent_grads, by = "Major_code")

merge_df
```

### Remove multiple Column in R using dplyr (by name and index) -- TidyVerse

```{r}

merge_df <- select(merge_df, -c(Major_category.y, Major.y))

merge_df

```

### To remove a common suffix from column names in an R data frame

```{r}
colnames(merge_df)<-gsub(".x","",colnames(merge_df))
merge_df
```
### Quantiles 

To demonstrate this new flexibility in a more useful situation, let’s take a look at quantile(). quantile() was hard to use previously because it returns multiple values. Now it’s straightforward

```{r}
merge_df %>% 
  group_by(Major_category) %>% 
  summarise(x = quantile(Grad_unemployment_rate, c(0.25, 0.5, 0.75)), q = c(0.25, 0.5, 0.75))
```

### Summerizing the merge_df using 'summarise' function on 'Grad_unemployed' column

```{r}
summary1 <- merge_df %>% 
  group_by(Major_category) %>% 
  summarise(Grad_unemployed = mean(Grad_unemployed, na.rm = TRUE)) %>% 
  arrange(desc(Grad_unemployed))

summary1
```

### Visualizing the trend of Grad_unemployed vs Major_category

```{r}
ggplot(data = summary1) +
  geom_count(mapping = aes(x = Grad_unemployed, y = Major_category,  color = Major_category) )
```


