---
title: "New york Times Web APIs"
output: html_document
---


### Loading the Libraries:
```{r}
library(httr)
library(jsonlite)
library(dplyr)
```

## Brief Overview:

Task: In this assignment, the task is to choose one of the New York Times APIs, request the API key, 
contruct the URL, perform an API call to fetch the JSON data from the server, and finally 
transform the data into R dataframe.
Purpose: is to collect the current data information into the dataframe in R for analysis.
API Selected: In this assignment I am choosing the Books API, which provides information about book 
reviews and The New York Times Best Sellers list.

## Access the NYT API Key:

```{r}
apikey <- "04GA736x2mowjM9AOusr0XYdn2t3uVSN"
```

## Construct the URL for NYTimes API call
### The lists/names service returns a list of all the NYT Best Sellers Lists.
### Using "current" for {date} to get the latest list.

```{r}
url <-  paste0("https://api.nytimes.com/svc/books/v3//lists/current/hardcover-fiction.json?api-key=",apikey)

# url
```

## Fetch response:

```{r}
response <- GET(url)

response
```

## Checking the status code of the response

```{r}

response$status_code

```

## Extract headers with headers() function

```{r}
headers(response)
```

## Extract content from the body with $content or content()--use 'text'

```{r}
data <- content(response, "text")
```

## Making the data readable using "prettify" function

```{r}
json_data <- content(response, "text")

prettify(json_data)
```



## Converting into a R dataframe

```{r}
books_df <- fromJSON(url, flatten = TRUE)
books_df <- books_df$results
books_df <- books_df$books

books_df
```

## To reorder the column by position 

```{r}

books_df <- books_df[, c(2, 11, 12, 8, 7, 6, 9)]

books_df
```



## Check the data type

```{r}
str(books_df)
```

## Column names of the dataframe

```{r}
books_df %>%
  names
```

```{r}
books_df$rank_last_week
```

## sort dataframe by column

```{r}
 
books_df <- books_df[order(books_df$rank_last_week),]
books_df
```

## Filter the rows of our data with the == operator

```{r}

books_df[books_df$rank_last_week == 1, ]  

```

## Conclusion & Analysis:

The New York Times Books Review is a weekly paper in which the current fictional and non_fictional books are reviewed.
The Books API provides information about book reviews and The New York Times Best Sellers lists.
According to the data fetched through API call to New York Times, I found that the # 1 "rank_last_week"
book is "RUN, ROSE, RUN". The publisher of the book is "Little, Brown" and the primary_isbn13 is "9780759554344"
