---
title: "Week 10 Assignment – Sentiment Analysis"
output: html_document
---


## Re-create base analysis:

Re-create and analyze primary code from the textbook.
Provide citation to text book, using a standard citation syntax like APA or MLA.

### Import Libraries

```{r}
library(tidytext)
library(dplyr)
```

Emily Dickinson wrote this text:

```{r}
text <- c("Because I could not stop for Death -",
          "He kindly stopped for me -",
          "The Carriage held but just Ourselves -",
          "and Immortality")
```

Turn text into a tidy text dataset, we first need to put it into a data frame.
```{r}
library(dplyr)
text_df <- tibble(line = 1:4, text = text)

text_df
```

To both break the text into individual tokens (a process called tokenization) and 
transform it to a tidy data structure we need to import Library tidytext 

```{r}
library(tidytext)

text_df %>%
  unnest_tokens(word, text)
```
### Tidying the works of Jane Austen
```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]",
                                           ignore_case = TRUE)))) %>%  ungroup()
 

original_books
```

```{r}
library(tidytext)
tidy_books <- original_books %>%
  unnest_tokens(word, text)

tidy_books
```

