---
title: "Week 10 Assignment – Text Mining"
output: html_document
---


### Assigned Task:

Re-create and analyze primary code from the textbook.
Provide citation to text book, using a standard citation syntax like APA or MLA.


### Citation to text book, using a standard citation MLA syntax.

### Silge and Robinson - O'Reilly - 2017

---

### Import Libraries

```{r}
library(tidytext)
library(dplyr)
```

Emily Dickinson wrote this text:

```{r}
text <- c("Because I could not stop for Death -",
          "He kindly stopped for me -",
          "The Carriage held but just Ourselves -",
          "and Immortality")
```

Turn text into a tidy text dataset, we first need to put it into a data frame.
```{r}
library(dplyr)
text_df <- tibble(line = 1:4, text = text)

text_df
```

To both break the text into individual tokens (a process called tokenization) and 
transform it to a tidy data structure we need to import Library tidytext 

```{r}
library(tidytext)

text_df %>%
  unnest_tokens(word, text)
```
Tidying the works of Jane Austen, creating line number, and chapter column 
in the dataframe
```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]",
                                           ignore_case = TRUE)))) %>%  ungroup()
 

original_books
```

```{r}
library(tidytext)

tidy_books <- original_books %>%
  unnest_tokens(word, text)

tidy_books
```
We can remove stop words (kept in the tidytext dataset stop_words) with an anti_join().
```{r}
data(stop_words)

tidy_books <- tidy_books %>%
  anti_join(stop_words)

tidy_books

```
use dplyr’s count() to find the most common words in all the books as a whole
```{r}
tidy_books %>%
  count(word, sort = TRUE) 
```

Using tidy tools, our word counts are stored in a tidy data frame. 
This allows us to pipe this directly to the ggplot2 package.

```{r}
library(ggplot2)

tidy_books %>%
  count(word, sort = TRUE) %>%
  filter(n > 600) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```
### The gutenbergr package
```{r}
library(gutenbergr)

hgwells <- gutenberg_download(c(35, 36, 5230, 159))

hgwells
```

```{r}
tidy_hgwells <- hgwells %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)
```

The most common words in these novels of H.G. Wells

```{r}
tidy_hgwells %>%
  count(word, sort = TRUE)
```
We will again use the Project Gutenberg ID numbers for each novel and access the texts using gutenberg_download()
```{r}
bronte <- gutenberg_download(c(1260, 768, 969, 9182, 767))
```

```{r}
tidy_bronte <- bronte %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)
```
The most common words in these novels of the Brontë sisters
 “time”, “eyes”, and “hand” are in the top 10 for both H.G. Wells and the Brontë sisters.
```{r}
tidy_bronte %>%
  count(word, sort = TRUE)

```
To calculate the frequency for each word for the works of Jane Austen, the Brontë sisters, and H.G. Wells by binding the data frames together. We can use pivot_wider() and pivot_longer() from tidyr to reshape our dataframe 
```{r}
library(tidyr)

frequency <- bind_rows(mutate(tidy_bronte, author = "Brontë Sisters"),
                       mutate(tidy_hgwells, author = "H.G. Wells"), 
                       mutate(tidy_books, author = "Jane Austen")) %>% 
  mutate(word = str_extract(word, "[a-z']+")) %>%
  count(author, word) %>%
  group_by(author) %>%
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  pivot_wider(names_from = author, values_from = proportion) %>%
  pivot_longer(`Brontë Sisters`:`H.G. Wells`,
               names_to = "author", values_to = "proportion")

frequency

```

We use str_extract() here because the UTF-8 encoded texts from Project Gutenberg have some examples of words with underscores around them to indicate emphasis (like italics). The tokenizer treated these as words, but we don’t want to count “_any_” separately from “any” as we saw in our initial data exploration before choosing to use str_extract()


Comparing the word frequencies of Jane Austen, the Brontë sisters, and H.G. Wells
```{r}
library(scales)

# expect a warning about rows with missing values being removed
ggplot(frequency, aes(x = proportion, y = `Jane Austen`, 
                      color = abs(`Jane Austen` - proportion))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), 
                       low = "darkslategray4", high = "gray75") +
  facet_wrap(~author, ncol = 2) +
  theme(legend.position="none") +
  labs(y = "Jane Austen", x = NULL)
```

Let’s quantify how similar and different these sets of word frequencies are using a correlation test. How correlated are the word frequencies between Austen and the Brontë sisters, and between Austen and Wells?

```{r}
cor.test(data = frequency[frequency$author == "Brontë Sisters",],
         ~ proportion + `Jane Austen`)
```

```{r}
cor.test(data = frequency[frequency$author == "H.G. Wells",], 
         ~ proportion + `Jane Austen`)
```
We conclude in the plots, the word frequencies are more correlated between the Austen and Brontë novels than between Austen and H.G. Wells.
