---
title: "Week 7 - Working with HTML, JSON, and XML file"
output: html_document
---


# Import Libraries:
```{r}
library(rvest)
library(dplyr)
library(tidyverse)
library(DT)
```

# Reading HTML file
```{r }
# Get working directory path
github_link <- "https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/week7/books_data.html"

github_link
```

```{r}
# Start by reading a HTML page with read_html():

html_data<- read_html(github_link)

html_data
```
# Converting the HTML data into dataframe
```{r}
data <-html_table(html_data)
html_df <- as.data.frame(data)

html_df
```

# Load the JSON Package


```{r}
library(jsonlite)
```

# Read a JSON file into a dataframe

```{r}
# Give the input file name to the function.
github_link = "https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/week7/books_data.json"



json_df <- read_json(github_link, simplifyVector = TRUE)

json_df
```

# loading the library and other important packages

```{r}
library("XML")
library("methods")

```

# the contents of books_data.xml are parsed

```{r}

github_link = "https://raw.githubusercontent.com/uzmabb182/CUNY-SPS-Assignments/main/data_607/week7/books_data.XML"

github_link
```

```{r}
library(httr)
library(XML)

r = GET(github_link)
xml_data <- xmlTreeParse(r, useInternal=TRUE)
 
print(xml_data)
```


```{r}

xml_df <- xmlToDataFrame(xml_data)
xml_df
```

# Now to compare these html and json dataframes, simply pass them to the comparedf() function:

# Please find the conclusion in the very end of the file
```{r}
library(arsenal)

```

```{r}
comparedf(html_df, json_df)
```

# Use summary() to get a more detailed summary

```{r}
summary(comparedf(html_df, json_df))
```
# Now compare html and xml dataframes
```{r}
comparedf(html_df, xml_df)
```
# Use summary() to get a more detailed summary

```{r}
summary(comparedf(html_df, xml_df))
```
# Now compare json and xml dataframes
```{r}
comparedf(json_df, xml_df)
```

# Use summary() to get a more detailed summary

```{r}
summary(comparedf(json_df, xml_df))
```
# Conclusion:

### Comparing all the datafarmes, I conclude that all the dataframes are equal

```{r}
# The data frames from the HTML and XML route are the same
html_df == xml_df
```

```{r}
# The data frames from the HTML and JSON route are the same
html_df == json_df
```

```{r}
# The data frames from the HTML and JSON route are the same
xml_df == json_df
```


